### 2015-01-16 preamble.tex
initial commit: skeletal draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  PREAMBLE: sets up compiler modes, loads packages, defines macros, etc
%  Steve Rodney, 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
% COMPILER MODES
%%%%%%%%%%%%%%%%%%%%

%% Set default compiler modes 
%% (these will be overridden by the -options.tex  file below, if it exists)

% Manuscript mode : 
%  True  : single-column, double-spaced, markup-ready style (with comments)
%  False : print in two-col single-space publication-ready style (no comments)
\newif\ifms
\msfalse

% Grayscale  mode : use grayscale figures when available
\newif{\ifgrayscale}
\grayscalefalse

% Changetext mode : highlight modified text in bold blue font
\newif{\ifchangetext}
\changetextfalse

%% Read in the -options.tex file (generated by the Makefile)
%%  to set the compile-mode options
\InputIfFileExists{\jobname-options}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% manuscript mode settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifms
  \documentclass[manuscript]{aastex}
  \usepackage{natbib,amsmath,verbatim}
  \received{}
  \revised{}
  \accepted{}
  \citestyle{aa}
  \newcommand{\cmt}[1]{\textcolor{red}{#1}}   % Red text  for printing comments 
\else
  %\documentclass[12pt,ms,iop]{emulateapj}
  \documentclass[revtex4,iop]{emulateapj}
  \font\sevenrm=cmr7 
  \citestyle{aa} 
  \newcommand{\cmt}[1]{}
  %\slugcomment{submitted to ApJ} 
  %\journalinfo{astro-ph/xxxxxx}  % top left corner of first page.
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% changetext  mode settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifchangetext
  % Changed text is highlighted in bold, blue font 
  \newcommand{\change}[1]{{\bf \textcolor{blue}{#1}}}
\else
  % Changed text is indistinguishable
  \newcommand{\change}[1]{#1}
\fi

%%%%%%%%%%%%%%%%%%%%
% PACKAGES INCLUDED
%%%%%%%%%%%%%%%%%%%%
\usepackage{natbib}   % reference citations and bibliography
\usepackage{amsmath}  % extended math symbols
\usepackage{verbatim} % verbatim text formatting
\usepackage{enumerate}% enumerated lists
\usepackage{amssymb}  % extended symbols lib
%\usepackage{url}      % url text formatting
\usepackage{hyperref}      % url text formatting
\usepackage[usenames]{color}  % colored text
%\usepackage{multirow}  % muti-row table cells
%\usepackage{amsmath}  % extended equations lib (split)
%\usepackage{mathrsfs} % extended math fonts (mathscr)
%\usepackage{paralist} % inline enumeration (for Table ref lists)
%\usepackage{authblk}
\usepackage{appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDF mode settings : Auto-select eps or pdf figures 
% based  on the compiler used (i.e. latex vs pdflatex)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifx\pdfoutput\undefined
  \pdffalse
  \DeclareGraphicsExtensions{.eps,.ps}
\else
  \ifnum\pdfoutput=1
    % \pdftrue
    \DeclareGraphicsExtensions{.pdf,.png,.jpg}
  \else
    % \pdffalse
    \DeclareGraphicsExtensions{.eps,.ps}
  \fi
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHOR-DEFINED MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper aliases 
\defcitealias{Patel:2014}{P14}
\def\P14{\citetalias{Patel:2014}}

\def\tomas{HFF14Tom}

% General purpose usefulness:
\newcommand{\etal}{{et al.~}}                                             
\def\eg{{e.g.}}
\def\ie{{i.e.}}
\def\etc{{etc.}}
\newcommand{\lta}{\lesssim}                                               
\newcommand{\gta}{\gtrsim}                                                
\newcommand{\gt}{\gtsim}
\newcommand{\kms}{\,\rm km\,s^{-1}}

%\def\pmstat{ \ensuremath{ \substack{ \resizebox{0.7em}{0.5em}{$\pm$} \\ \resizebox{0.7em}{0.2em}{stat}} } }
\def\pmstat{\ensuremath{\substack{\pm \\ \mbox{\scalebox{0.45}{stat}} } } }
\def\pmsyst{\ensuremath{\substack{\pm \\ \mbox{\scalebox{0.45}{syst}} } } }

% Cosmology:
\def\Om{\ensuremath{\Omega_{\rm m}}}
\def\Ot{\ensuremath{\Omega_{\rm tot}}}
\def\Ob{\ensuremath{\Omega_{\rm b}}}
\def\OL{\ensuremath{\Omega_{\Lambda}}}
\def\Ok{\ensuremath{\Omega_{\rm k}}}
\def\om{\ensuremath{\omega_{\rm m}}}
\def\ob{\ensuremath{\omega_{\rm b}}}
\def\wo{\ensuremath{w_0}}
\def\wa{\ensuremath{w_{\rm a}}}
\def\lcdm{$\Lambda$CDM}
\def\LCDM{$\Lambda$CDM}
\def\wcdm{$w$CDM}
\def\Ho{\ensuremath{H_0}}
\def\DA{\ensuremath{D_A}}
\def\DL{\ensuremath{D_L}}

% Astronomy:
\def\arcsec{\ensuremath{^{\prime\prime}}} 
\def\kms{\ensuremath{{\rm km s}^{-1}}}
\def\hgpcq{\mbox{$h^{-3}$Gpc$^3$}}
\def\hmpcq{\mbox{$h^{-3}$Mpc$^3$}}
\def\perhmpcq{\mbox{$h^{3}$Mpc$^{-3}$}}
\def\hmpc{\mbox{$h^{-1}$Mpc}}
\def\hmpci{\mbox{$h$\,Mpc$^{-1}$}}
\def\mpc{\mbox{Mpc}}
\def\mpci{\mbox{Mpc$^{-1}$}}
\def\mpcq{\mbox{Mpc$^{-3}$}}
\def\Msun{\mbox{M$_{\odot}$}}
\def\Av{\mbox{$A_V$}}
\def\Rv{\mbox{$R_V$}}

% Supernovae : 
\newcommand{\CCSN}{CC\,SN}
\newcommand{\CCSNe}{CC\,SNe}
\newcommand{\TNSN}{TN\,SN}
\newcommand{\TNSNe}{TN\,SNe}
\newcommand{\SNIa}{SN\,Ia}
\newcommand{\SNeIa}{SNe\,Ia}
\newcommand{\SNRz}{SNR($z$)}
\def\Mch{\mbox{M$_{\rm Ch}$}}
\def\Ni{\ensuremath{^{56}\mbox{Ni}}}
\newcommand{\dmfifteen}{\ensuremath{\Delta\mbox{m}_{15}}}
\newcommand{\deltamfifteen}{\ensuremath{\Delta\mbox{m}_{15}}}

% Missions:
\def\Hubble{{\it Hubble}}
\def\Hubbles{{\it Hubble's}}
\def\Spitzer{{\it Spitzer}}
\def\Chandra{{\it Chandra}}
\def\Herschel{{\it Herschel}}
\def\XMM{{\it XMM}}

% Institutions
\newcommand{\JHU}{Department of Physics and Astronomy, The Johns Hopkins University, Baltimore, MD 21218.}
\newcommand{\elsewhere}{St. Elsewhere University.}
\newcommand{\STScI}{Space Telescope Science Institute, Baltimore, MD 21218.}
\newcommand{\Berkeley}{Department of Astronomy, University of California, Berkeley, CA 94720-3411.}
\newcommand{\Riverside}{Department of Physics and Astronomy, University of California, Riverside, CA 92521.}
\newcommand{\WKU}{Department of Physics, Western Kentucky University, Bowling Green, KY 42101.}
\newcommand{\Copenhagen}{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, DK-2100 Copenhagen, Denmark.}
\newcommand{\Arizona}{Department of Astronomy, University of Arizona, Tucson, AZ 85721.}
\newcommand{\SantaCruz}{Department of Astronomy and Astrophysics, University of California, Santa Cruz, CA 92064.}
\newcommand{\NotreDame}{Department of Physics, University of Notre Dame, Notre Dame, IN 46556.}
\newcommand{\TelAviv}{Department of Astrophysics, Tel Aviv University, 69978 Tel Aviv, Israel.}
\newcommand{\Rutgers}{Department of Physics and Astronomy, Rutgers, The State University of New Jersey, Piscataway, NJ 08854.}
\newcommand{\CfA}{Harvard/Smithsonian Center for Astrophysics, Cambridge, MA 02138.}
\newcommand{\Minnesota}{Department of Astronomy, University of Minnesota, 116 Church Street SE, Minneapolis, MN 55455, USA.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Page Setup 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifms
\else
  \renewcommand{\topfraction}{0.9}
  \renewcommand{\bottomfraction}{0.9}
  \renewcommand{\textfraction}{0.1}
  \renewcommand{\floatpagefraction}{0.9}
  \renewcommand{\dbltopfraction}{0.9}
  \renewcommand{\dblfloatpagefraction}{0.9}
\fi