### 2015-04-03 authors.tex
updated authors list

\author{Steven~A.~Rodney\altaffilmark{1,2}}

\author{Brandon Patel}

\author{Daniel Scolnic}

\author{Ryan Foley}

\author{Alberto Molino}

\author{Marusa Bradac}

\author{Gabriel Brammer}

\author{Jose~Maria~Diego}

\author{Jens Hjorth}

\author{Austin Hoag}

\author{Mathilde Jauzac}

\author{Saurabh Jha}

\author{Patrick Kelly}

\author{Daniel Lam}

\author{Julian~Merten}

\author{Johan Richard}

\author{Keren Sharon}

\author{Louis-Gregory Strolger}

\author{Tommaso Treu}

\author{Adi Zitrin}

\altaffiltext{1}{\JHU}
\altaffiltext{1}{\Rutgers}
\altaffiltext{1}{\Chicago}
\altaffiltext{1}{\Illinois}
\altaffiltext{1}{\CSIC}
\altaffiltext{1}{\SaoPaulo}
\altaffiltext{1}{\UCDavis}
\altaffiltext{1}{\STScI}
\altaffiltext{1}{\Copenhagen}
\altaffiltext{1}{\UCDavis}
\altaffiltext{1}{\Rutgers}
\altaffiltext{1}{\Berkeley}
\altaffiltext{1}{\STScI}
\altaffiltext{1}{\WKU}
\altaffiltext{1}{\UCLA}
\altaffiltext{1}{\CalTech}
\altaffiltext{2}{Hubble fellow}