### 2015-04-05 authors.tex
added System 3.3 fig and Intro, updated authors.tex, updated mag fig and lensingtest fig

\author{Steven~A.~Rodney\altaffilmark{1,*}}
\author{Brandon Patel\altaffilmark{2}}
\author{Daniel Scolnic\altaffilmark{3}}
\author{Ryan Foley\altaffilmark{4}}
\author{Alberto Molino\altaffilmark{5,6}}
\author{Mathilde Jauzac\altaffilmark{7,8}}
\author{Gabriel Brammer\altaffilmark{9}}
\author{Marusa Bradac\altaffilmark{10}}
\author{Jose~Maria~Diego\altaffilmark{11}}
\author{Jens Hjorth\altaffilmark{12}}
\author{Or Graur\altaffilmark{13,14}}
\author{Austin Hoag\altaffilmark{10}}
\author{Saurabh Jha\altaffilmark{2}}
\author{Patrick Kelly\altaffilmark{15}}
\author{Daniel Lam\altaffilmark{16}}
\author{Elinor Medezinski\altaffilmark{1,17}}
\author{Julian~Merten\altaffilmark{18}}
\author{Johan Richard\altaffilmark{19}}
\author{Keren Sharon\altaffilmark{20}}
\author{Louis-Gregory Strolger\altaffilmark{9,21}}
\author{Tommaso Treu\altaffilmark{22}}
\author{Liliya Williams\altaffilmark{23}}
\author{Adi Zitrin\altaffilmark{24,*}}

\altaffiltext{1}{\JHU}
\altaffiltext{2}{\Rutgers}
\altaffiltext{3}{\Chicago}
\altaffiltext{4}{\Illinois}
\altaffiltext{5}{\CSIC}
\altaffiltext{6}{\SaoPaulo}
\altaffiltext{7}{\Durham}
\altaffiltext{8}{\KwaZulu}
\altaffiltext{9}{\STScI}
\altaffiltext{10}{\UCDavis}
\altaffiltext{11}{\IFCA}
\altaffiltext{12}{\Copenhagen}
\altaffiltext{13}{\NYU}
\altaffiltext{14}{\AMNH}
\altaffiltext{15}{\Berkeley}
\altaffiltext{16}{\HongKong}
\altaffiltext{17}{\Hebrew}
\altaffiltext{18}{\Oxford}
\altaffiltext{19}{\CRAL}
\altaffiltext{20}{\Michigan}
\altaffiltext{21}{\WKU}
\altaffiltext{22}{\UCLA}
\altaffiltext{23}{\Minnesota}
\altaffiltext{24}{\CalTech}
\altaffiltext{*}{Hubble fellow}