### 2015-04-07 TABLE/magnifications.tex
Incorporated comments from KS, MJ, AZ   * added positions   * added a section on cosmological parameter uncertainty   * subsumed the section 3.3 discussion into the preceding subsection

\begin{deluxetable*}{lcccp{1.4in}p{2.1in}}
\tablecolumns{5}
\tablecaption{Predicted magnifications for SN \tomas\ from lens models.\label{tab:PredictedMagnifications}}
\tablehead{ \colhead{Model} & \colhead{Best\tablenotemark{a}} & \colhead{Median\tablenotemark{b}} & \colhead{68\% Conf. Range\tablenotemark{b}} & \colhead{References} & \colhead{Description}}
\startdata
Sharon(v1)   & 2.53     & 2.57  &   2.41$-$2.75 & \citealt{Jullo:2007}  & LENSTOOL\tablenotemark{c} parametric, strong-lensing based model\\
Sharon(v2)   & 2.73     & 2.69  &   2.63$-$2.83 &   \citealt{Jullo:2007};\citealt{Johnson:2014} & LENSTOOL parametric, strong-lensing based model.\\
CATS-SL      & 2.25     & 2.27  &   2.23$-$2.32 &   \citealt{Jullo:2007,Richard:2014} &  CATS\tablenotemark{d} team implementation of LENSTOOL parametric strong-lensing based model.\\
CATS-SL+WL   & \nodata  & 2.62  &   2.44$-$2.80 &   \citealt{Jullo:2007,Richard:2014} &  CATS team LENSTOOL parametric model with both strong and weak lensing constraints.\\
CATS-SL(v2)  & 3.37     & 3.35  &   3.18$-$3.53 &   \citealt{Richard:2014,Jauzac:2014} & Updated version of the CATS-SL model, adds 33 new multiply-imaged galaxies.\\
GLAFIC       & 2.32     & 2.28  &   2.17$-$2.35 &   \citealt{Oguri:2010,Ishigaki:2015} & Parametric strong-lensing model using the {\tt GLAFIC} code.\tablenotemark{e} \\
Zitrin-NFW   & 2.07     & 2.27  &   2.05$-$2.50 &   \citealt{Zitrin:2013a} &  Parametric strong-lensing model using PIEMD\tablenotemark{f} profiles for galaxies and NFW\tablenotemark{g} profiles for dark matter halos.\\
Zitrin-LTM   & 2.64     & 2.96  &   2.58$-$3.73 &   \citealt{Zitrin:2009a} & Parametric strong-lensing model, adopts the Light-Traces-Mass assumption for both the luminous and dark matter.\\
Bradac(v1)   & 3.15     & 2.45  &   2.29$-$2.64 &   \citealt{Bradac:2005,Bradac:2009} & {\tt SWUnited} : Free-form, strong+weak-lensing based model. Errors from bootstrap resampling only weak-lensing constraints.\\
Bradac(v2)   & 2.21     & 2.23  &   2.20$-$2.28 &   \citealt{Wang:2015} & \change{Updated version of the {\tt SWUnited} model with new strong-lensing constraints from HFF imaging. Errors from bootstrap resampling only strong-lensing constraints.}\\
Williams     & 2.67     & 2.78  &   1.64$-$5.46 &   \citealt{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014} & {\tt GRALE}\tablenotemark{h} : Free-form strong-lensing model using a genetic algorithm.  \\
Merten       & 2.31     & 2.22  &   2.03$-$2.89 &   \citealt{Merten:2009,Merten:2011} &  {\tt SaWLENS},\tablenotemark{i} Grid-based free-form strong+weak lensing based model using adaptive mesh refinement.\\
Lam			 & \nodata  & 2.77  &   2.41$-$3.13 &   \citealt{Sendra:2014,Lam:2014} & {\tt WSLAP+}\tablenotemark{j} : Free-form model strong-lensing using a grid-based method, supplemented by deflections fixed to cluster member galaxies.\\
Diego\tablenotemark{k}   & \nodata  & 2.10  &   1.74$-$2.46 &   \citealt{Sendra:2014,Lam:2014} & Alternative implementation of the {\tt WSLAP+}\tablenotemark{j} model, using a different set of strong-lensing constraints and redshifts.
\enddata
\tablenotetext{a}{The magnification returned for the optimal version of each model, as independently defined by each lens modeling team.}  
\tablenotetext{b}{Median magnification from 100-600 Monte Carlo realizations of the model. Confidence ranges enclose 68\%\ of the realized values.}  
\tablenotetext{c}{{\tt LENSTOOL} : \url{http://projects.lam.fr/repos/lenstool/wiki}}
\tablenotetext{d}{{\tt CATS} : Clusters As TelescopeS lens modeling team.}
\tablenotetext{e}{{\tt GLAFIC} : \url{http://www.slac.stanford.edu/~oguri/glafic/}}
\tablenotetext{f}{PIEMD: Pseudo Isothermal Elliptical Mass Distrubition}
\tablenotetext{g}{NFW : Navarro-Frenk-White mass density profile \citep{Navarro:1997}.}
\tablenotetext{h}{{\tt GRALE} : GRAvitational LEnsing.}
\tablenotetext{i}{{\tt SaWLENS} : Strong and Weak LENSing analysis code. \url{http://www.julianmerten.net/codes.html}}
\tablenotetext{j}{{\tt WSLAP+} : Weak and Strong Lensing Analysis Package plus member galaxies \change{(Note: no weak-lensing constraints used for Abell 2744)}}
\tablenotetext{k}{No uncertainty estimates were available for the Diego implementation of the {\tt WSLAP+} model, so we adopt the uncertainties from the closely related Lam model.}
\end{deluxetable*}