### 2015-04-07 TABLE/positions.tex

\begin{deluxetable}{lllll}
\tablecolumns{5}
\tablecaption{Coordinates\tablenotemark{a} of \tomas, host, and cluster. \label{tab:Positions}}
\tablehead{ 
    \colhead{Object}
  & \colhead{R.A.}
  & \colhead{Decl.}
  & \colhead{R.A.}
  & \colhead{Decl.}\\
    \colhead{}
  & \colhead{(h:m:s)}
  & \colhead{(d:m:s)}
  & \colhead{(deg)}
  & \colhead{(deg)}
}
\startdata
\tomas       &  00:14:17.87 &  -30:23:59.7  &  3.574458 &  -30.399917 \\
Host galaxy  &  00:14:17.88 &  -30:24:00.6  &  3.574483 &  -30.400175 \\
Abell 2744   &  00:14:21.20 &  -30:23:50.1  &  3.588333 &  -30.397250
\enddata
\tablenotetext{a}{Coordinates in Right Ascension and Declination, on the J2000 reference system.}
\end{deluxetable}